---
title: "65 World Indexes"
author: "Pranav Pandya | s_pandya@stud.hwr-berlin.de"
date: "26th February 2017"
output: html_document
---

-----------------------------------------------------------

### Exploring "65 World Indexes" dataset in R with ggplot

```{r, message=FALSE, warning=FALSE}
#Making csv file readable (separating by comma)
#WI = read.csv("HumanDevelopmentIndex.csv", sep=";", stringsAsFactors = FALSE, dec=",")

WI = read.csv("Kaggle.csv")
#names(WI)

library(ggplot2) 
library(ggrepel)
library(dplyr)
library(scales)

WI$Human.Development.Index.HDI.2014 = as.numeric(WI$Human.Development.Index.HDI.2014)
WI$Human.Development.Index.HDI.2014 <- round(WI$Human.Development.Index.HDI.2014, 2)

HDI = WI[,1:2]
names(HDI) <- c("country", "HDI")

HDIplot= ggplot(HDI)+
                geom_point(aes(country, HDI), size = 4, color = 'grey') +
                geom_label_repel(
                  aes(country, HDI, fill = factor(HDI), label = country),
                  fontface = 'bold', color = 'white',
                  box.padding = unit(0.35, "lines"),
                  point.padding = unit(0.5, "lines"),
                  segment.color = 'grey50')+
                  ggtitle("Human Development Index by County")+
                  labs(x="Countries",y="Human Development Index (2014)") + 
                  theme(axis.text.x = element_text(angle = 90))

```

###1.0 Human development index
```{r, message=FALSE, warning=FALSE}
WI$Human.Development.Index.HDI.2014 = as.numeric(WI$Human.Development.Index.HDI.2014)
WI$Human.Development.Index.HDI.2014 <- round(WI$Human.Development.Index.HDI.2014, 2)

HDI = WI[,1:2]
names(HDI) <- c("country", "HDI")

HDIplot= ggplot(HDI)+
                geom_point(aes(country, HDI), size = 4, color = 'grey') +
                geom_label_repel(
                  aes(country, HDI, fill = factor(HDI), label = country),
                  fontface = 'bold', color = 'white',
                  box.padding = unit(0.35, "lines"),
                  point.padding = unit(0.5, "lines"),
                  segment.color = 'grey50')+
                  ggtitle("Human Development Index by County")+
                  labs(x="Countries",y="Human Development Index (2014)") + 
                  theme(axis.text.x = element_text(angle = 90))
```

###1.0 Dot plot illustrating Human development index by country
```{r, message=FALSE, warning=FALSE, fig.width= 15, fig.height= 14}
HDIplot
```

- Almost 60% countries have HDI less than 0.65

###2.0 Co2 emission

```{r, message=FALSE, warning=FALSE}

co2 = WI[, c("Id", "Carbon.dioxide.emissionsAverage.annual.growth")]
names(co2) <- c("country", "co2emission")

co2$co2emission = as.numeric(co2$co2emission)
co2$co2emission <- round(co2$co2emission, 2)

co2plot = ggplot(co2, aes(country, co2emission)) +
            ggtitle("Avg. Annual growth in Co2 emission by Country")+
            labs(x="Countries",y="Avg annual Co2 emission") +
            geom_point(stat = "identity", color = 'grey', aes(fill = factor(co2emission))) + 
            geom_label_repel( 
              aes(country, co2emission, fill = factor(co2emission), label = country), 
              fontface = 'bold', color = 'white',
              box.padding = unit(0.35, "lines"),
              point.padding = unit(0.5, "lines"),
              segment.color = 'grey50') +
            geom_text(aes(label = ifelse(co2emission>5, (co2emission), '')), size = 4, color = 'gray51') +
            theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))

```

###2.1 Dot plot illustrating annual growth in Co2 emission by countries
```{r, message=FALSE, warning=FALSE, fig.width= 15, fig.height= 14}
co2plot 
```

- Bhutan, Equitorial Guinea and Oman represents shockingly high average annual growth in Co2 emission compared to other countries. 
- However based on geographical area, China (6.53) and Indonesia (6.65) have considerably high Co2 emission.

###3.0
- to be continued...